Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-17 concentration_argument.tex
updated sections 5 and 6 with respect to changes in section 3

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$, $k_n$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and let $\Kcal_{C}(\ell_n)$ be defined as in~\eqref{eq:def_K_C_set}. In addition, define $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ is any function satisfying 
\[
	\Mu{\BallPo{0,y}}(1 - \phi_n^{(1)}(y)) \le \hat{\mu}_n(y) \le \Mu{\BallPo{0,y}}(1 + \phi_n^{(2)}(y))
\]
where $\phi_n^{(i)} : \R_+ \to \R_+$ are such that $\sup_{0 \le y \le (1-\varepsilon)R_n} \phi_n(y)^{(i)} = \smallO{1}$ for any $0 < \varepsilon < 1$ and $\partial \phi_n^{(i)}(y)/\partial y = \bigT{\phi_n^{(i)}(y)}$. Then the following holds for some $C > 0$ large enough:
\begin{enumerate}
\item for any sequence of functions $h_n(y) : \R_+ \mapsto \R$ such that $h_n(y) = \hat{\rho}_n(y,k_n) \bigO{k_n^s e^{\beta y}}$ for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Kcal_{C}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\item If, in addition, $h_n(y) = \hat{\rho}_n(y,k_n) g_n(y)$ with $g_n$ satisfying $g_n(y(1+\smallO{1}) = (1+\smallO{1})g_n(y)$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Rcal_n} \rho(y,k_n) g_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\end{enumerate}
\end{lemma}

@@

The proof of the second statement follows the same line of reasoning as above. First note that by Lemma~\ref{lem:concentration_argument} and the first statement, it is enough to show that
\[
	\int_{a_n^-}^{a_n^+} \rho_n(y,k_n) g_n(y) e^{-\alpha y} \dd y
	= \left(1 + o(1)\right) \int_{a_n^-}^{a_n^+} \rho(z,k_n) g_n(z) e^{-\alpha z} \dd z.
\] 
For this we observe that~\eqref{eq:concentration_argument_z_upper_bound} and~\eqref{eq:concentration_argument_z_lower_bound} imply that $y = z(y) + \smallO{1}$ for $a_n^- \le y \le a_n^+$ and hence, by our assumption on $g_n(y)$ we have that
\[
	g_n(y) = g_n(z + \smallO{1}) = (1 + \smallO{1})g_n(z).
\]
Therefore, with the change of variables~\eqref{eq:concentration_argument_variable_change} we get
\begin{align*}
	\int_{a_n^-}^{a_n^+} \rho_n(y,k_n) g_n(y) e^{-\alpha y} \dd y
	&=  \int_{a_n^-}^{a_n^+} \Prob{\Po(\hat{\mu}_n(y)) = k_n} g_n(y) e^{-\alpha y} \dd y\\
	&= (1 + \smallO{1})\int_{a_n^-}^{a_n^+}  \Prob{\Po(\mu(z))}
		g_n(z) e^{-\alpha z} \frac{\hat{\mu}_n^\prime(z)}{\mu^{\prime}(z)} \dd z\\
	&= \left(1 + o(1)\right) \int_{a_n^-}^{a_n^+} \rho(z,k_n) g_n(z) e^{-\alpha z} \dd z.
\end{align*}

\end{proof}

>>>>

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$, $k_n$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and let $\Kcal_{C}(\ell_n)$ be defined as in~\eqref{eq:def_K_C_set}. In addition, define $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ is any function satisfying 
\[
	\Mu{\BallPo{0,y}}(1 - \phi_n^{(1)}(y)) \le \hat{\mu}_n(y) \le \Mu{\BallPo{0,y}}(1 + \phi_n^{(2)}(y))
\]
where $\phi_n^{(i)} : \R_+ \to \R_+$ are such that $\sup_{0 \le y \le (1-\varepsilon)R_n} \phi_n(y)^{(i)} = \smallO{1}$ for any $0 < \varepsilon < 1$ and $\partial \phi_n^{(i)}(y)/\partial y = \bigT{\phi_n^{(i)}(y)}$. Then the following holds for some $C > 0$ large enough:
\begin{enumerate}
\item for any sequence of functions $h_n(y) : \R_+ \mapsto \R$ such that $h_n(y) = \hat{\rho}_n(y,k_n) \bigO{k_n^s e^{\beta y}}$ for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Kcal_{C}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\item for any uniformly bounded function $h : \R_+ \mapsto \R$,
\[
	\int_{\Rcal_n} h(y) \hat{\rho}_n(y,k_n) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Rcal_n} \rho(y,k_n) h(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\end{enumerate}
\end{lemma}

@@

The proof of the second statement follows the same line of reasoning as above. First note that by Lemma~\ref{lem:concentration_argument} and the first statement, it is enough to show that
\[
	\int_{a_n^-}^{a_n^+} \rho_n(y,k_n) h(y) e^{-\alpha y} \dd y
	= \left(1 + o(1)\right) \int_{a_n^-}^{a_n^+} \rho(z,k_n) h(z) e^{-\alpha z} \dd z.
\] 
For this we observe that~\eqref{eq:concentration_argument_z_upper_bound} and~\eqref{eq:concentration_argument_z_lower_bound} imply that there exists a $b_n \to 0$ such that $y = z(y) + b_n$, uniformly for $a_n^- \le y \le a_n^+$. Hence, if we define $g_n(z) = h(z + b_n)$, then $g_n(z)$ is uniformly bounded and converges point wise to $h(z)$. Therefore, with the change of variables~\eqref{eq:concentration_argument_variable_change} we get
\begin{align*}
	\int_{a_n^-}^{a_n^+} \rho_n(y,k_n) h(y) e^{-\alpha y} \dd y
	&=  \int_{a_n^-}^{a_n^+} \Prob{\Po(\hat{\mu}_n(y)) = k_n} h(y) e^{-\alpha y} \dd y\\
	&= (1 + \smallO{1})\int_{a_n^-}^{a_n^+}  \Prob{\Po(\mu(z))}
		g_n(z) e^{-\alpha z} \frac{\hat{\mu}_n^\prime(z)}{\mu^{\prime}(z)} \dd z\\
	&= (1 + \smallO{1})\int_{a_n^-}^{a_n^+}  \Prob{\Po(\mu(z))}
		g_n(z) e^{-\alpha z} \dd z\\
	&= \left(1 + o(1)\right) \int_{a_n^-}^{a_n^+} \rho(z,k_n) h(z) e^{-\alpha z} \dd z.
\end{align*}
where the last line follow by dominated convergence.
\end{proof}

### 2018-11-17 proof_overview.tex
updated sections 5 and 6 with respect to changes in section 3

\begin{proposition}\label{prop:clustering_ast_H_Pois}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n\ge 1}$ be any increasing sequence such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, as $n \to \infty$,
\[
	\Exp{\left|c_{\H,n}^\ast(k_n) - c_{\HP,n}^\ast(k_n)\right|} = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}
\MS{I think we can drop the condition to be increasing here and below, also the $n^{1/3}$ should be removed.}

@@

\begin{proposition}[Coupling result for local clustering]\label{prop:couling_c_H_P}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n\ge 1}$ be any increasing sequence such that $k_n = \smallO{\min\left\{n^{\frac{1}{2\alpha + 1}}, n^{\frac{1}{3}}\right\}}$. Then, as $n \to \infty$,
\[
	\Exp{\left|c_{\HP,n}^\ast(k_n) - c_{\Pcal,n}^\ast(k_n)\right|} = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}

>>>>

\begin{proposition}\label{prop:clustering_ast_H_Pois}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n\ge 1}$ be any increasing sequence such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, as $n \to \infty$,
\[
	\Exp{\left|c_{\H,n}^\ast(k_n) - c_{\HP,n}^\ast(k_n)\right|} = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}
%\MS{I think we can drop the condition to be increasing here and below, also the $n^{1/3}$ should be removed.}

@@

\begin{proposition}[Coupling result for local clustering]\label{prop:couling_c_H_P}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n\ge 1}$ be any positive sequence such that $k_n \ge 3$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, as $n \to \infty$,
\[
	\Exp{\left|c_{\HP,n}^\ast(k_n) - c_{\Pcal,n}^\ast(k_n)\right|} = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}
